On the Utility of Azimuth and Decentralized Identity

### 2024-08-29 mss.tex
Add refs

\section{Alternative Options for the \textsc{pki}}

@@

What is the best way to enable this portability (another shared principle/design goal for \textsc{ssi}s/\textsc{did}s) while maintaining global consistency? The reader who is familiar with Nockchain will foresee the punchline:  zero-knowledge proofs.

@@

\subsection{Groundwire, comets & Bitcoin ordinals}

We also note that an early design consideration of the 21e8 Bitcoin Ordinals project was to be able to host an Azimuth-like \textsc{pki} system. Recent debates in the ecosystem about the Azimuth PKI, namely those between Zenith and Mayflower, have instigated discussion about alternative ways of handling identiy across the Urbit network. Groundwire is a proposal to experiment with both new PKI governance models, and to make use of the comet-level address space in a more robust way. Implementation details are still forthcoming, but inital plans are to try to prototype a Gall agent that could support a PKI on Bitcoin that would allow comets to rotate their keys (updating the current 'self-attestation' model), and make use of different non-hierarchical models for routing and peer-discovery. Additionally, instead of using the finite 'higher level' address space for spam resistance, this proposal would use mechanisms like L1 Bitcoin fee costs, or burning and timelocking coins.

As with the Zenith and Mayflower proposals, the Groundwire proposal makes statements about the governance and fundrasing directions of the Urbit project. These topics are out of scope for the focus of this article, but deserve ongoing discussion in the appropriate forums.

>>>>

\section[{Alternative Options for the \textsc{pki}}]{Alternative Options for the PKI}

@@

What is the best way to enable this portability (another shared principle/design goal for \textsc{ssi}s/\textsc{did}s) while maintaining global consistency? The reader who is familiar with Nockchain will anticipate the punchline:  zero-knowledge proofs.

@@

\subsection{GroundWire, comets, and Bitcoin Ordinals}

An early design consideration of the 21e8 Bitcoin Ordinals project \citep{Ordinals} was reportedly to be able to host an Azimuth-like \textsc{pki} system, and the desire to tie Urbit's fortunes more closely to those of Bitcoin persists in segments of the Urbit community. A recent debate throughout the ecosystem about the Azimuth \textsc{pki}, namely between Zenith and Mayflower, has instigated a wide-ranging discussion about alternative ways of handling identiy across the Urbit network. GroundWire proposes to experiment with both new \textsc{pki} governance models, and to make use of the comet-level address space in a more robust way \citep{Hamilton2024}. Implementation details are still forthcoming, but the initial plan is to prototype a Gall agent to support a \textsc{pki} on Bitcoin that would allow comets to rotate their keys (updating the current `self-attestation' model), and make use of different non-hierarchical models for routing and peer-discovery. Additionally, instead of using the finite `higher level' address space for spam resistance, this proposal would use mechanisms like L1 Bitcoin fee costs, or burning and timelocking coins.

As with the Zenith and Mayflower proposals, the GroundWire proposal makes statements about the governance and fundrasing directions of the Urbit project. These topics are out of scope for the focus of this article, but merit ongoing discussion in the appropriate forums.
